# Data Processing

The goal is to compare  rainfall and discharge data in order to compute the following metrics: 1) total quantity of response flow, 2) magnitude of peak response discharge, 3) duration of response, 4) lag to peak time (start time of storm to time of peak discharge), and 5) runoff ratio (depth of Q (mm) divided by the depth of rainfall (mm) for the rain storm associated with the runoff response). 


## Precipitation

Before starting this analysis, I identified discrete rainfall events from the tipping bucket rain gauge output using the USDA Rainfall Intensity Summarization Tool (RIST; [@ARS2013]). RIST calculated the following metrics for each event: storm duration (hrs), rain depth (mm); maximum intensity (mm h^−1^) over 5-, 15-, 30-, and 60-min intervals; and erosivity. Because the stage data is recorded every 15 minutes, I'm using the start time of the maximum intensity 15 minute interval as the storm start time. The function below formats the RIST output into the form I want.

```{r, include=FALSE}
library(EcoHydRology)
library(tidyverse)
library(lubridate)
library(plotly)
library(kableExtra)
library(ggthemes)
```

``` {r, show_col_types = FALSE}
#function to read in rain files
rain_reader <- function(rain_file){
  df <-read.delim(rain_file, sep = "", header = TRUE) %>%
  select(-c(1,3:7))
}

#function to process the storm output 
rain_format <- function(rain_file){
  df<- rain_file %>%
  select(Max_15_start_date, Max_15_start_time, Precip.in.) %>%
  rename(date =1, time=2, p_in =3) %>%
  mutate(datetime = paste(date, time, sep = " ")) %>%
  filter(!(p_in ==0)) %>%
  mutate(p_mm = p_in*25.4) %>%
  mutate(DateTime = mdy_hms(datetime, tz = "GMT"))  
}

#read in and format michigan ditch and bl4 rain data
mich_rain <- rain_reader("raw_sitedata/michiganditch/michigan_storm.txt")
bl4_rain <- rain_reader("raw_sitedata/bl4/bl4_storm.csv")

mich_storm <- rain_format(mich_rain)
bl4_storm <- rain_format(bl4_rain)
```


## Stream Stage 

Michigan Ditch stream stage was recorded by a pressure transducer that doesn't take into account atmospheric pressure, so the data needs to be corrected using barometic data. The continuous stage data is then calibrated according to the manual stage measurements.

```{r michigan, warning=F, fig.cap='Michigan Ditch stream stage. The red dots indicate manual stage measurements.', out.width='80%', fig.asp=.75, fig.align='center'}
#read in pt data
michiganditch_stage_pt=read_csv('raw_sitedata/michiganditch/michiganditch_stage_composite_pt.csv', show_col_types = FALSE) %>%
  rename(Pw_kPa = 3)
michiganditch_stage_pt$DateTime = as_datetime(michiganditch_stage_pt$datetime,tz="GMT", format="%m/%d/%Y %H:%M")

#read in baro data
michiganditch_baro=read_csv('raw_sitedata/michiganditch/michiganditch_baro_composite.csv', show_col_types = FALSE) %>%
  filter(!is.na(Pressure_kPa)) %>%
  rename(Pa_kPa = 2)
michiganditch_baro$DateTime = as_datetime(michiganditch_baro$datetime,tz="GMT", format="%m/%d/%Y %H:%M")

#join pt and baro and convert to stage in cm
mich_stage = full_join(michiganditch_baro, michiganditch_stage_pt, by="DateTime",all=TRUE) %>%
  mutate(stage_kPa = Pw_kPa - Pa_kPa) %>%
  mutate(stage_cm = (stage_kPa*101.97162129779)/10) %>%
  select(DateTime, stage_cm)

#read in file with manual stage measurements
discharge_michiganditch=read_csv('raw_sitedata/michiganditch/discharge_michiganditch.csv', show_col_types = FALSE)
discharge_michiganditch$DateTime = as.POSIXct(discharge_michiganditch$datetime,tz="GMT",format="%m/%d/%Y %H")

#merge the manual discharge measurements with the sensor time series
michditch_stage=full_join(mich_stage,discharge_michiganditch,by='DateTime') %>%
  select(1,2,7,9) 

#correct stage based on manual stage measurements
michditch_stage<-mutate(michditch_stage, stage_corr_cm = stage_cm)

#offset adjust
michditch_stage$stage_corr_cm=ifelse(michditch_stage$DateTime >ymd_hms('2021-06-21 10:30:00'),
                                     michditch_stage$stage_corr_cm-1,michditch_stage$stage_corr_cm)

michditch_stage$stage_corr_cm=ifelse(michditch_stage$DateTime>ymd_hms('2021-08-05 11:00:00'),
                                     michditch_stage$stage_corr_cm-2.3,michditch_stage$stage_corr_cm)

#delete sensor spikes
michditch_stage <- michditch_stage %>%
  filter(!(DateTime < ymd_hms("2021-01-10 09:30:00") |
             DateTime == ymd_hms("2021-06-21 09:45:00") |
             stage_corr_cm <= 0))

#plot the sensor and manual stage time series
michiganditch=ggplot()+
  geom_line(data=michditch_stage,aes(x=DateTime,y=stage_corr_cm))+
  geom_point(data=michditch_stage, aes(x=DateTime,y=manual_stage_cm),color='red')+
  labs(x= "Time", y= "Stage (cm)") +
  theme_bw()
  
ggplotly(michiganditch)
```


## Discharge


```{r}

```


## Chapters and sub-chapters

There are two steps to cross-reference any heading:

1. Label the heading: `# Hello world {#nice-label}`. 
    - Leave the label off if you like the automated heading generated based on your heading title: for example, `# Hello world` = `# Hello world {#hello-world}`.
    - To label an un-numbered heading, use: `# Hello world {-#nice-label}` or `{# Hello world .unnumbered}`.

1. Next, reference the labeled heading anywhere in the text using `\@ref(nice-label)`; for example, please see Chapter \@ref(cross). 
    - If you prefer text as the link instead of a numbered reference use: [any text you want can go here](#cross).

## Captioned figures and tables

Figures and tables *with captions* can also be cross-referenced from elsewhere in your book using `\@ref(fig:chunk-label)` and `\@ref(tab:chunk-label)`, respectively.

See Figure \@ref(fig:nice-fig).

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center', fig.alt='Plot with connected points showing that vapor pressure of mercury increases exponentially as temperature increases.'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Don't miss Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(pressure, 10), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```
